---
title: "Data Visualization Final Project: \nHealth Insurance Coverage in the United States"
author: "Grace Kong, Iris Chen & Fang Liu (Group P)"
output:
  html_document: default
  html_notebook: default
---

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
rm(list=ls())

library(leaflet)
library(maps)
library(rgdal)
library(DT)
library(ggplot2)
library(ggthemes)
library(plotly)
library(magrittr)
library(readxl)
library(plyr)
library(dplyr)
library(tidyr)
library(readr)
library(stringi)
library(RColorBrewer)
library(countrycode)
require(gridExtra)
```


```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## IMPORT DATA

# General ACA information
aca_general_raw <- read_excel("Data/states_aca_general.xlsx", sheet=2, col_names=FALSE)

# States health outcomes
health_ind_raw <- read_excel("Data/states_health_ind.xlsx", sheet=4, col_names=FALSE)

# Insurance coverage data
  # Long format
insurance_long <- read_excel("Data/insurance-clean.xlsx", sheet=1, col_names=TRUE)
  # Wide format
insurance <- read_excel("Data/insurance-clean.xlsx", sheet=2, col_names=TRUE)

# Demographic information
demographics1 <- read.csv("Data/Population_Age_Income.csv", header = TRUE)
  # Provided with Assignment 4 (want the yearly population data)
demographics2 <- read_excel("Data/PopulationEstimates.xls")

# Election Results
election <- read_excel("Data/US_Presidential_Results_by_State_1828-2016.xlsx", sheet=2, col_names=FALSE)
```

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## CLEAN DATA

# List of states abbreviations
state_abbreviations <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY")


## CLEAN INSURANCE RATES DATA (WIDE)

# Drop unnecessary columns
insurance <- insurance[1:50, -1]

# Rename variables
colnames(insurance) <- c("state_abb", "state", "uninsured_num_2008", "uninsured_pct_2008", "uninsured_num_2009", "uninsured_pct_2009", "uninsured_num_2010", "uninsured_pct_2010", "uninsured_num_2011", "uninsured_pct_2011", "uninsured_num_2012", "uninsured_pct_2012", "uninsured_num_2013", "uninsured_pct_2013", "uninsured_num_2014", "uninsured_pct_2014", "uninsured_num_2015", "uninsured_pct_2015")

# Create variables for insurance rates
insurance$insured_pct_2008 <- 100 - insurance$uninsured_pct_2008
insurance$insured_pct_2009 <- 100 - insurance$uninsured_pct_2009
insurance$insured_pct_2010 <- 100 - insurance$uninsured_pct_2010
insurance$insured_pct_2011 <- 100 - insurance$uninsured_pct_2011
insurance$insured_pct_2012 <- 100 - insurance$uninsured_pct_2012
insurance$insured_pct_2013 <- 100 - insurance$uninsured_pct_2013
insurance$insured_pct_2014 <- 100 - insurance$uninsured_pct_2014
insurance$insured_pct_2015 <- 100 - insurance$uninsured_pct_2015

# Reorder variables
insurance <- insurance[ , c("state_abb", "state", "uninsured_num_2008", "uninsured_pct_2008", "insured_pct_2008", "uninsured_num_2009", "uninsured_pct_2009", "insured_pct_2009", "uninsured_num_2010", "uninsured_pct_2010", "insured_pct_2010", "uninsured_num_2011", "uninsured_pct_2011", "insured_pct_2011", "uninsured_num_2012", "uninsured_pct_2012", "insured_pct_2012", "uninsured_num_2013", "uninsured_pct_2013", "insured_pct_2013", "uninsured_num_2014", "uninsured_pct_2014", "insured_pct_2014", "uninsured_num_2015", "uninsured_pct_2015", "insured_pct_2015")]


## CLEAN INSURANCE RATES DATA (LONG)

# Rename variables
colnames(insurance_long) <- c("year", "state", "population", "uninsured_num", "uninsured_pct", "insured_num", "insured_pct")

# Add state abbreviations
  # First create a look-up table (based on wide data)
state_abb_lookup <- insurance[ , c("state", "state_abb")]
  # Then do the matching
insurance_long$state_abb <- state_abb_lookup$state_abb[match(insurance_long$state, state_abb_lookup$state)]
  # Reorder variable
insurance_long <- insurance_long[ , c(1:2, ncol(insurance_long), 3:(ncol(insurance_long)-1))]

# Order observations (by state, then by year)
insurance_long <- insurance_long[order(insurance_long$state, insurance_long$year), ]


## CLEAN ACA GENERAL DATA

# Keep only relevant rows and columns
aca_general <- aca_general_raw[6:57, 1:73]

# Label variables for ACA general data
names(aca_general) <- c("state", "unins_all_pct_10", "unins_all_pct_15", "unins_all_decr_pct", "unins_all_decr", "cov_emp", "cov_parents_plan", "lifetime_lim_preACA_tot", "lifetime_lim_preACA_child", "lifetime_lim_preACA_adultM", "lifetime_lim_preACA_adultF", "cov_private_tot", "cov_private_child", "cov_private_adultM", "cov_private_adultF", "premium_emp_avg_growth_pct_00to10", "premium_emp_avg_growth_pct_10to15", "premium_emp_savings_15", "premium_emp_savings_16", "MLR_rebate_beneficiaries_12", "MLR_rebate_amt_12", "MLR_rebate_beneficiaries_13", "MLR_rebate_amt_13", "MLR_rebate_beneficiaries_14", "MLR_rebate_amt_14", "MLR_rebate_beneficiaries_15", "MLR_rebate_amt_15", "MLR_rebate_amt_12to15", "medicaid_enroll_13", "medicaid_enroll_16", "medicaid_enroll_child_13", "medicaid_enroll_incr_13to16", "medicaid_full_duals", "medicaid_partial_duals", "medicaid_full_or_partial_duals", "state_has_expanded", "insurance_incr_medicaid", "cholest_scr_incr_medicaid", "mammogram_incr_medicaid", "papsmear_incr_medicaid", "clinic_care_incr_medicaid", "all_care_incr_medicaid", "phycisian_visit_ann_incr_medicaid", "depression_decr_medicaid", "good_health_incr_medicaid", "deaths_ann_decr_medicaid", "catastrophic_oop_ann_decr_medicaid", "indebted_ppl_decr_medicaid", "fed_spending_net_incr_inMil", "uncompensated_care_decr_inMil", "mental_substance_elig_medicaid_share", "mental_substance_elig_medicaid", "preexisting_condition_09", "cov_mkt_plan_16", "cov_mkt_16", "receive_tax_credit", "avg_tax_credit", "receive_cost_sharing", "avg_num_mkt_plans_avail_17", "cov_mkt_under75D_pct", "cov_mkt_under100D_pct", "cov_offmkt_elig_tax_credit", "rate_review_funds_to_state", "HIECP_grant_award_to_state", "medicare_enroll_16", "medicare_benef_donuthole", "medicare_benef_donuthole_savings_tot", "medicare_benef_donuthole_savings_avg", "medicare_partB_free_prev_services", "medicare_partB_free_prev_services_share", "medicare_incr_readmit_rate", "medicare_avoided_readmit", "accountable_care_org_num")

# Trim white space on state variable
aca_general$state <- stri_trim_both(aca_general$state)

# Change format of variables
  # Apart from T/F variable for "state has expanded"
aca_general$state_has_expanded[aca_general$state_has_expanded=="yes"] <- TRUE 
aca_general$state_has_expanded[aca_general$state_has_expanded=="no"] <- FALSE 
  # Convert others to numeric form
for (j in c(2:35, 37:73)) {
  aca_general[[j]] <- as.numeric(aca_general[[j]])
}

# Save USA data separately
aca_general_USA <- aca_general[aca_general$state=="United States", ]

# Drop USA and DC data
aca_general <- aca_general[!(aca_general$state %in% c("United States", "District of Columbia")), ]

# Add state abbreviations
aca_general$state_abb <- state_abbreviations
aca_general <- aca_general[ , c(1, ncol(aca_general), 2:(ncol(aca_general)-1))]

# Drop raw data
rm(aca_general_raw)


## CLEAN HEALTH INDICATOR DATA

# Keep only relevant rows and columns
health_ind <- health_ind_raw[4:6703, c(1, 3:6)]

# Label variables
names(health_ind) <- c("state", "measure", "year", "rate", "rank")

# Simplify date (years) - only keep last year if range was given
health_ind$year[health_ind$year=="07/2009 - 06/2012"] <- "2012"
health_ind$year[health_ind$year=="07/2010 - 06/2013"] <- "2013"
health_ind$year[health_ind$year=="07/2012 - 06/2015"] <- "2015"
health_ind$year[health_ind$year=="10/2013-9/2014"] <- "2014"
health_ind$year[health_ind$year=="2008-09"] <- "2009"
health_ind$year[health_ind$year=="2010-11"] <- "2011"
health_ind$year[health_ind$year=="2011-12"] <- "2012"
health_ind$year[health_ind$year=="2011/12"] <- "2012"
health_ind$year[health_ind$year=="2012-13"] <- "2013"
health_ind$year[health_ind$year=="2013-14"] <- "2014"
health_ind$year[health_ind$year=="2013(Q2-Q4)"] <- "2013"
health_ind$year[health_ind$year=="2014-15"] <- "2015"
health_ind$year[health_ind$year=="2015(Q2-Q4)"] <- "2015"

# Export to excel to manually label variable names
unique_measures <- unique(as.factor(health_ind$measure))
  # Copied in from excel sheet
unique_measures_relabelled <- c("a.summary_access", "a.unins_adult", "a.unins_child", "a.no_care_bc_cost_adult", "a.high_OOP_relative_under65", "a.at_risk_no_routine_doc_adult", "a.no_dental_adult", "q.summary_prev_treat", "q.with_usual_care_adult", "q.with_cancer_screening_adult", "q.with_vaccines_adult", "q.with_medical_home_child", "q.with_prev_medical_dental_child", "q.with_mental_healthcare_child", "q.with_vaccines_infant", "q.drug_should_avoid_medicare", "q.drug_should_avoid_3conditions_medicare", "q.good_health_provider_medicare", "q.mortality_4conditions_medicare", "q.CLABSI_infection_ratio", "q.info_recovery_hospitalized", "q.good_hospital_staff_hospitalized", "q.improve_mobility_homehealth", "q.improved_wounds_homehealth", "q.sores_NHres", "q.antipsychotic_med_NHres", "u.summary_avoidable_hosp_cost", "u.hosp_asthma_child", "u.hosp_ambulatory_65to74yrs", "u.hosp_ambulatory_above75yrs", "u.30day_hosp_readmit_medicare", "u.30day_hosp_readmit_NHres", "u.hosp_6mos_NHres", "u.hosp_medicare_homehealth", "u.avoidable_ER_medicare", "u.tot_reimb_employer_ins", "u.tot_reimb_medicare", "h.summary_healthy_lives", "h.deaths_amenable", "h.yrs_lost_potential_life_before75", "h.deaths_breast_cancer_F", "h.deaths_colorectal_cancer", "h.deaths_suicide", "h.deaths_infant_mortality", "h.poor_health_adult", "h.smoke_adult", "h.obese_adult", "h.obese_child", "h.poor_dental_adult", "u.premium_emp_private", "u.premium_emp_private_unadj", "u.reimb_medicare_unadj", "u.deaths_amenable_black", "u.deaths_amenable_white", NA, "q.with_prev_screening_above50yrs", NA)

# Replace variable names with abbreviated version
for (i in 1:57) {
  health_ind$measure <- replace(health_ind$measure, health_ind$measure==unique_measures[i], unique_measures_relabelled[i])
}

# Remove the NA's
health_ind <- health_ind[!is.na(health_ind$measure), ]

# Remove duplicates
  # By state, measure.year
health_ind <- health_ind[!duplicated(health_ind[ , 1:3]), ]

# Interact measure and year
health_ind$measure_year <- interaction(health_ind$measure, health_ind$year)

# Convert to character variable
health_ind$measure_year <- as.character(health_ind$measure_year)

# Pre-reshape: Store existing data as "long"
health_ind_long <- health_ind

# Reshape wide
health_ind <- as.data.frame(cbind(health_ind$state, health_ind$measure_year, health_ind$rate))
  # Drop rank data for now
names(health_ind) <- c("state", "measure_year", "rate")
health_ind <- reshape(health_ind,
                        idvar = "state",
                        timevar = "measure_year",
                        direction = "wide")

# Drop "rate" from the name
names(health_ind)[2:ncol(health_ind)] <- substr(names(health_ind), 6, nchar(names(health_ind)))[2:ncol(health_ind)]

# Drop variables who have all entries as NA
  # Due to 'unbalanced panel' in reshaping
  # First create a vector to record if variable is all NA
variable_all_NA <- vector(mode="logical", length=ncol(health_ind))
for (j in 1:ncol(health_ind)) {
  if(sum(is.na(health_ind[[j]]))==52) {
    variable_all_NA[j] <- TRUE
  }
}
  # Only keep the variables that are NOT all NA
health_ind <- health_ind[ , !variable_all_NA]

# Change format of variables
  # To character variable
health_ind$state <- as.character(health_ind$state)
  # To numeric variable
for (j in c(2:164)) {
  health_ind[[j]] <- as.character(health_ind[[j]])
  health_ind[[j]] <- as.numeric(health_ind[[j]])
}

# Save USA data separately
health_ind_USA <- health_ind[health_ind$state=="United States", ]

# Drop USA and DC data
health_ind <- health_ind[!(health_ind$state %in% c("United States", "District of Columbia")), ]

# Label state abbreviations
health_ind$state_abb <- state_abbreviations
health_ind <- health_ind[ , c(1, ncol(health_ind), 2:(ncol(health_ind)-1))]

# Drop raw data
rm(health_ind_raw)


## CLEAN DEMOGRAPHIC (1) DATA

# Drop unnecessary row and column
demographics1 <- demographics1[-1,-1]

# Rename variables
colnames(demographics1) <- c("state","income","population", "ppl_age0to4","ppl_age5to9","ppl_age10to14","ppl_age15to19","ppl_age20to24", "ppl_age25to29","ppl_age30to34","ppl_age35to39","ppl_age40to44", "ppl_age45to49","ppl_age50to54","ppl_age55to59","ppl_age60to64", "ppl_age65to69","ppl_age70to74","ppl_age75to79","ppl_age80to84","ppl_age85plus")

# Drop observations DC and PR
demographics1 <- demographics1[!demographics1$state=="District of Columbia" & !demographics1$state=="Puerto Rico", ]

# Label state abbreviations
demographics1$state_abb <- state_abbreviations
demographics1 <- demographics1[ , c(1, ncol(demographics1), 2:(ncol(demographics1)-1))]

# Change format of variables
  # To character variable
demographics1$state <- as.character(demographics1$state)
  # To numeric variable
for (j in c(3:ncol(demographics1))) {
  demographics1[[j]] <- as.character(demographics1[[j]])
  demographics1[[j]] <- as.numeric(demographics1[[j]])
}

# BEA regions
demographics1$BEA_region <- ""
demographics1$BEA_region[demographics1$state_abb %in% c("ME", "NH", "VT", "MA", "CT", "RI")] <- "New England"
demographics1$BEA_region[demographics1$state_abb %in% c("NY", "NJ", "PA", "MD", "DE")] <- "Mideast"
demographics1$BEA_region[demographics1$state_abb %in% c("WI", "IL", "MI", "IN", "OH")] <- "Great Lakes"
demographics1$BEA_region[demographics1$state_abb %in% c("WV", "KY", "VA", "TN", "NC", "SC", "AR", "LA", "MS", "AL", "GA", "FL")] <- "Southeast"
demographics1$BEA_region[demographics1$state_abb %in% c("ND", "SD", "NE", "KS", "MN", "IA", "MO")] <- "Plains"
demographics1$BEA_region[demographics1$state_abb %in% c("MT", "ID", "WY", "UT", "CO")] <- "Rocky Mountains"
demographics1$BEA_region[demographics1$state_abb %in% c("AZ", "NM", "TX", "OK")] <- "Southwest"
demographics1$BEA_region[demographics1$state_abb %in% c("WA", "OR", "CA", "NV", "AK", "HI")] <- "Far West"

# Tag income level (category)
demographics1$income_level <- ""
demographics1$income_level[demographics1$income >= 60000] <- "High"
demographics1$income_level[demographics1$income >= 50000 & demographics1$income < 60000] <- "Upper Middle"
demographics1$income_level[demographics1$income >= 45000 & demographics1$income < 50000] <- "Lower Middle"
demographics1$income_level[demographics1$income < 45000] <- "Low"
  # Convert to factor variable
demographics1$income_level <- factor(demographics1$income_level, levels = c("Low", "Lower Middle", "Upper Middle", "High"))


## CLEAN DEMOGRAPHIC (2) DATA

colnames(demographics2) <- demographics2[2, ]

# Drop first few rows (non-observations)
demographics2 <- demographics2[-(1:2), ]

# Drop county-level data,  keep state-level data only
demographics2 <- demographics2[demographics2$Area_Name == "United States" | demographics2$State != lag(demographics2$State), ]

# Keep only columns of population from 2010 to 2015
demographics2 <- demographics2[ , c("State", "Area_Name", "POP_ESTIMATE_2010", "POP_ESTIMATE_2011", "POP_ESTIMATE_2012", "POP_ESTIMATE_2013", "POP_ESTIMATE_2014", "POP_ESTIMATE_2015")]

# Change format of variables
  # To character variable
for (j in c(3:8)) {
  demographics2[[j]] <- as.numeric(demographics2[[j]])
}

# Rename variables
colnames(demographics2) <- c("state_abb", "state", "population_2010", "population_2011", "population_2012", "population_2013", "population_2014", "population_2015")


## CLEAN ELECTION DATA

# Keep only 2012 and 2008 data, and data for the 50 states
election <- election[ 4:53 , c("X0", "X5", "X6", "X9", "X10")]

# Rename variables
colnames(election) <- c("state", "Dem_pct_2012", "Rep_pct_2012", "Dem_pct_2008", "Rep_pct_2008")

# Add state abbreviations
  # Match with look-up table
election$state_abb <- state_abb_lookup$state_abb[match(election$state, state_abb_lookup$state)]

# Change format of variables
  # To numeric variable
for (j in c(4:6)) {
  election[[j]] <- as.numeric(election[[j]])
}

# Calculate winning party for each state
election$party_2012 <- ifelse(election$Dem_pct_2012 > election$Rep_pct_2012, "Democratic", "Republican")
election$party_2008 <- ifelse(election$Dem_pct_2008 > election$Rep_pct_2008, "Democratic", "Republican")

# Reorder variables
election <- election[ , c("state", "state_abb", "Dem_pct_2012", "Rep_pct_2012", "party_2012", "Dem_pct_2008", "Rep_pct_2008", "party_2008")]
```


# Health Insurance Coverage

__Figure 1__

Click the link to access our Shiny App on [Uninsured Rates in the United States from 2008-2015](https://gracekongyx.shinyapps.io/health_insurance_shiny/).
   

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## SOME DATA PREPARATION

# Look-up information into insurance and insurance_long, on:
  # Party (2012)
insurance_long$party_2012 <- election$party_2012[match(insurance_long$state, election$state)]
insurance$party_2012 <- election$party_2012[match(insurance$state, election$state)]
  # BEA region
insurance_long$BEA_region <- demographics1$BEA_region[match(insurance_long$state, demographics1$state)]
insurance$BEA_region <- demographics1$BEA_region[match(insurance$state, demographics1$state)]
  # Income level
insurance_long$income_level <- demographics1$income_level[match(insurance_long$state, demographics1$state)]
insurance$income_level <- demographics1$income_level[match(insurance$state, demographics1$state)]

# Also find nation-wide trend for insurance
insurance_long_US <- data.frame("year" = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015), "uninsured_pct" = c(16.8, 17.5, 18.2, 17.2, 16.9, 16.6, 13.3, 10.5))
```

__Figure 2A__
   
The following figure displays the trend in the uninsured rate between 2008 and 2015. Results are shown for the USA average (in black), as well as for every individual state (in color). The states are group by Bureau of Economic Analysis (BEA) regions - this allows us to analyze regional trends, and also to identify a state by filtering down to its respective region on this interactive plot.

We see a clear downward trend in the uninsured rate once most of the Affordable Care Act (ACA) provisions were implemented in 2014. The overall US uninsured rate decreased from 18.2% in 2010 to 10.5% in 2015.

However, note that the disparities in uninsured rates largely continue to persist among the states, with regional patterns. States of the Southeast and Southwest consistently had the highest uninsured rates, while states of New England consistenely had the lowest insured or highest health insurance coverage rates.
   
```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 2A

plot2a <- (ggplot(insurance_long, aes(year, uninsured_pct)) 
          + geom_line(aes(color = BEA_region, group = state, label = uninsured_num))
          + scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015))
          + scale_y_continuous(breaks = c(5, 10, 15, 20, 25, 30))
          + labs(x = "Year", y = "% Uninsured", color = "Region")
          + geom_line(data = insurance_long_US, lwd = 0.8)
          + geom_vline(xintercept = 2010, linetype = 2, color = "gray30")
          + geom_vline(xintercept = 2014, linetype = 2, color = "gray30")
          + annotate("text",x = 2009.4, y = 25, label = "Obama \nsigns ACA", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2013.4, y = 24.5, label = "Most ACA \nprovisions \ntake effect", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2014.2, y = 11, label = "USA \nAverage", color = "black",fontface = 2, size = 3)
          + ggtitle("Change in Percentage Uninsured by State, \nfrom 2008 to 2015")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0, size = 12)))
ggplotly(plot2a, dynamicTicks = FALSE, tooltip = c("state", "year", "uninsured_pct", "uninsured_num"))
```

__Figure 2B__

The following plot provides a visual depiction of the narrowing in dispersal of the states' uninsured rates over the period of 2008 to 2015. It can be seen that by 2015, there is a concentration of states' uninsured rates especially within the range of 7.5% to 12.5%.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 2B
plot2b <- (ggplot(insurance_long, aes(year, uninsured_pct))
          + geom_violin(scale = "area", aes(group = year), fill = "dark blue", color = NA)
          + scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015))
          + scale_y_continuous(breaks = c(5, 10, 15, 20, 25, 30))
          + labs(x = "Year", y = "% Uninsured")
          + ggtitle("Narrowing of Uninsured Rates Across States \nfrom 2008 to 2015")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.5, size = 12)))
plot2b
```
   
__Figure 3__

In the following plot, we explore the difference between Democratic and Republican states, based on the 2012 US Presidential Election results. Generally, the democratic states consistently have had largely have a lower uninsured rate than the Republican states, though both groups experienced declines in the uninsured rates in 2014. The rate of decline of uninsured rates in 2014 for some Republican states appeared to exceed that of the Democratic states, mainly because of their higher starting point. 

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 3
plot3 <- (ggplot(insurance_long, aes(year, uninsured_pct)) 
          + geom_line(aes(color = party_2012, group = state, label = uninsured_num))
          + scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015))
          + scale_y_continuous(breaks = c(5, 10, 15, 20, 25, 30))
          + scale_color_manual(values = c("steel blue", "firebrick1")) 
          + labs(x = "Year", y = "% Uninsured", color = "Party \n(2012 Winner)")
          + geom_line(data = insurance_long_US, lwd = 1)
          + geom_vline(xintercept = 2010, linetype = 2, color = "gray30")
          + geom_vline(xintercept = 2014, linetype = 2, color = "gray30")
          + annotate("text",x = 2009.4, y = 25, label = "Obama \nsigns ACA", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2013.4, y = 24.5, label = "Most ACA \nprovisions \ntake effect", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2014.2, y = 11, label = "USA \nAverage", color = "black",fontface = 2, size = 3)
          + ggtitle("Change in Percentage Uninsured by Party, \nfrom 2008 to 2015")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0, size = 12)))
ggplotly(plot3, dynamicTicks = FALSE,  tooltip = c("state", "year", "uninsured_pct", "uninsured_num"))
```

__Figure 4__

We also know that the Medicaid expansion was found to be a big driver of the fall in uninsured rates (accounting for more than half of the decrease). Hypothesizing that this was highly correlated to party affiliation, we made the following bar graph. It confirmed our suspicions that Democratic states were more likely to opt for Medicaid expansion, whereas Republican states were more likely to opt out. This could be a possible explanation for the continued lower insured rates in Democratic states after 2014.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
# Merge in Medicaid expansion data
insurance$expanded_medicaid <- aca_general$state_has_expanded[match(insurance$state, aca_general$state)]

plot4 <- (ggplot(insurance, aes(label = state_abb))
  + geom_bar(aes(x = party_2012, fill = expanded_medicaid), position = "dodge")
  + scale_fill_manual(values = brewer.pal(n = 2, name = "Dark2")[c(2,1)])
  + labs(x = "Party (2012 Winner)", y = "Number of States", fill = "Expanded \nMedicaid")
  + ggtitle("Medicaid Expansion and Party of State")
  + theme_minimal()
  + theme(plot.title = element_text(face="bold", hjust = 0.5, size = 12)))
plot4
```


__Figure 5__

We want to provide a clearer comparison across of the gains in insurance coverage due to the ACA. Thus, we plotted the decrease in uninsured rate (or equivalently, the increase in insured rate) for the states from 2010 to 2015. We find that it is the West Coast states of Oregon, Nevada and California that led in gains in insurance coverage. Republican states are not all ranked at the bottom, and many of them are in the middle of the pack (again due to their higher starting point).

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
insurance$unins_pct_decr_10_15 <- insurance$uninsured_pct_2010 - insurance$uninsured_pct_2015
insurance$unins_num_decr_10_15 <- insurance$uninsured_num_2010 - insurance$uninsured_num_2015

## PLOT 5
plot5 <- (ggplot(insurance, aes(x = reorder(state, unins_pct_decr_10_15), y = unins_pct_decr_10_15, fill = party_2012, label = unins_num_decr_10_15))
  + geom_col()
  + coord_flip()
  + labs(x = "", y = "Decrease in % Uninsured", fill = "Party \n(2012 Winner)")
  + scale_fill_manual(values = c("steel blue", "firebrick1")) 
  + ggtitle("Decrease in in Percentage Uninsured from 2010 to 2015")
  + theme_minimal()
  + theme(plot.title = element_text(face="bold", hjust = 0, size = 12), axis.text.y = element_text(size = 6.5))
)  
ggplotly(plot5, dynamicTicks = FALSE,  tooltip = c("unins_pct_decr_10_15", "unins_num_decr_10_15"))
```

__Figure 6A__

This plot shows insured rates in 2015 at a glance, looking at how they differ by party (2012 winner), BEA region and income level of the states. 

In terms of correlations, it appears that being a Democratic state is a strong predictor of a higher insurance rate. States in New England, the Mideast and the Great Lakes perform well in insurance rates, while states in the Southwest, Rocky Mountains and Southeast perform the poorest. As expected, high-income states have the highest insurance rates as more people would be able to afford insurance. Lower and lower-middle income states have lower coverage rates, but it is infact the lower-middle income states that have the lowest insurance rates.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 6A

# Subset relevant variables
insurance_2010_2015 <- insurance[ , c("state", "insured_pct_2010", "unins_pct_decr_10_15", "insured_pct_2015", "party_2012", "BEA_region", "income_level")]
names(insurance_2010_2015)[names(insurance_2010_2015)=="unins_pct_decr_10_15"] <- "ins_pct_incr_10_15"

# Boxplots by party
plot6a1 <- (ggplot(insurance_2010_2015, aes(x = reorder(party_2012, insured_pct_2015), y = insured_pct_2015)) 
          + geom_boxplot(aes(fill = party_2012, label = insured_pct_2015), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = c("steel blue", "firebrick"))
          + ggtitle("Democratic vs. Republican States")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.37, size = 12), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_blank()))

# Boxplots by BEA region
plot6a2 <- (ggplot(insurance_2010_2015, aes(x = reorder(BEA_region, insured_pct_2015), y = insured_pct_2015)) 
          + geom_boxplot(aes(fill = BEA_region, label = insured_pct_2015), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = brewer.pal(n = 8, name = "Dark2"))
          + ggtitle("By Bureau of Economic Analysis (BEA) Region")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.24, size = 12), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_blank()))

# Boxplots by income level
plot6a3 <- (ggplot(insurance_2010_2015, aes(x = income_level, y = insured_pct_2015)) 
          + geom_boxplot(aes(fill = income_level, label = insured_pct_2015), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = brewer.pal(n = 4, name = "Dark2"))
          + labs(x = "", y = "% Insured (2015)")
          + ggtitle("By State's Income Level")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.39, size = 12), legend.position = "none", axis.title.y = element_blank()))

# Combine plots vertically
grid.arrange(plot6a1, plot6a2, plot6a3, nrow = 3, top = "2015 Insurance Rates at a Glance")
```

__Figure 6B__

We now make these same categorizations, but looking at the gains in insurance coverage between 2010 and 2015. We notice that Republican states at actually have higher gains in insurance rate than Democratic states (although Democratic states have a higher 3rd quartile). The Far West, Southwest and Southeast make the greatest gains in insurance coverage, but still had the lowest insured rates in 2015. The lower the income of the state, the greater the gains in insurance coverage although they still end up with the lowest insurance rates in 2015.

Figures 6A and 6B together suggest that the trends by party, region and income have been preserved from 2010 to 2015, but the gap between the high-performing groups and the low-performing groups has narrowed.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 6B

# Boxplots by party
plot6b1 <- (ggplot(insurance_2010_2015, aes(x = reorder(party_2012, ins_pct_incr_10_15), y = ins_pct_incr_10_15)) 
          + geom_boxplot(aes(fill = party_2012, label = ins_pct_incr_10_15), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = c("steel blue", "firebrick"))
          + ggtitle("Democratic vs. Republican States")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.37, size = 12), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_blank()))

# Boxplots by BEA region
plot6b2 <- (ggplot(insurance_2010_2015, aes(x = reorder(BEA_region, ins_pct_incr_10_15), y = ins_pct_incr_10_15)) 
          + geom_boxplot(aes(fill = BEA_region, label = ins_pct_incr_10_15), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = brewer.pal(n = 8, name = "Dark2"))
          + ggtitle("By Bureau of Economic Analysis (BEA) Region")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.24, size = 12), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_blank()))

# Boxplots by income level
plot6b3 <- (ggplot(insurance_2010_2015, aes(x = income_level, y = ins_pct_incr_10_15)) 
          + geom_boxplot(aes(fill = income_level, label = ins_pct_incr_10_15), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = brewer.pal(n = 4, name = "Dark2"))
          + labs(x = "", y = "Increase in % Insured")
          + ggtitle("By State's Income Level")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.39, size = 12), legend.position = "none", axis.title.y = element_blank()))

# Combine plots vertically
grid.arrange(plot6b1, plot6b2, plot6b3, nrow = 3, top = "Increase in Insurance Rates (2010-2015) at a Glance")
```

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## SOME DATA RESHAPING

# Reshape aca_general (long), keeping the interesting quantities
  # Import uninsured number (2015)
aca_general$unins_all_15 <- insurance$uninsured_num_2015[match(aca_general$state, insurance$state)]
  # Keep only select variables
aca_general_long <- aca_general[ , c("state", "state_abb", "cov_emp", "cov_mkt_plan_16", "medicaid_enroll_16", "medicare_enroll_16", "unins_all_15")]
  # The actual reshaping
aca_general_long <- reshape(aca_general_long, 
                      varying = c("cov_emp", "cov_mkt_plan_16", "medicaid_enroll_16", "medicare_enroll_16", "unins_all_15"),
                      v.names = "Number",
                      timevar = "Source",
                      times = c("Employer", "Marketplace", "Medicaid / CHIP", "Medicare", "Uninsured"),
                      new.row.names = 1:1000,
                      direction = "long")

# Look-up information into aca_general_long, on:
  # Party (2012)
aca_general_long$party_2012 <- election$party_2012[match(aca_general_long$state, election$state)]
  # BEA region
aca_general_long$BEA_region <- demographics1$BEA_region[match(aca_general_long$state, demographics1$state)]
  # Income level
aca_general_long$income_level <- demographics1$income_level[match(aca_general_long$state, demographics1$state)]
  # Population
aca_general_long$population_2015 <- demographics2$population_2015[match(aca_general_long$state, demographics2$state)]

# Calculate percentage from number and population
aca_general_long$percentage <- aca_general_long$Number / aca_general_long$population_2015 * 100
```
  
  
__Figure 7__

Next, we investigate the main health insurance and coverage sources, namely employer coverage, the individual marketplace, Medicaid / Childrens' Health Insurance Program (CHIP), and Medicare for the aged. We broke down the results by party and income level, taking a simple (unweighted) average across the states in the categories. We observe that employer coverage rates increase as income level increases, especially for Democratic states. Democratic and low income states have more people on Medicaid / CHIP than Republican and low income states.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 7

aca_general_long_by_party_income <- ddply(aca_general_long, c("party_2012", "income_level", "Source"), summarise, avg_pct = mean(percentage))
aca_general_long_by_party_income$avg_pct <- round(aca_general_long_by_party_income$avg_pct, digits = 1)

plot7 <- (ggplot(aca_general_long_by_party_income, aes(income_level, avg_pct, fill = Source))
          + geom_col(position = "dodge") + facet_grid(party_2012 ~ .)
          + labs(x = "Income Level", y = "Average % in State", color = "Source of \nInsurance")
          + scale_fill_manual(values = brewer.pal(n = 6, name = "Dark2")[c(1:3,6,4)])
          + ggtitle("Sources of Health Insurance, \nby Party and Income Level")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0, size = 12)))
ggplotly(plot7, tooltip = c("income_level", "Source", "avg_pct"))
```
