---
title: "Data Visualization Final Project: \nHealth Insurance Coverage in the United States"
author: "Grace Kong, Iris Chen & Fang Liu (Group P)"
output:
  html_document: default
  html_notebook: default
---
     
# Introduction

## Goals of the project

In this project, we present a _state-level overview_ of the expansion of health insurance in the United States and its effects on health outcomes and cost outcomes, from 2008 to 2015 under President Obama's term. One of the hallmarks of Obama's presidency was the passage and implementation of the __Affordable Care Act (ACA)__ to increase health insurance coverage and regulate healthcare provision to protect consumers' rights. In addition, we are also interested in how people perceive the (ACA), as seen from articles related to the ACA from the New York Times.

__Research Questions:__

1) _Trends of health insurance expansion in the United States_: How has the percentage of uninsured Americans changed overall, and differentially by states, from 2008 to 2015? How has the composition of healthcare sources changed?

2) _Health outcomes_: How do the states compare in health outcomes, such as mortality rates and access to preventative and remedying treatments?

3) _Financial outcomes_: How do the states compare in financial and cost outcomes, such as the containment of out-of-pocket costs, insurance premium growth and government spending?

4) _Perceptions of the ACA_: We used API to download articles realted to ACA on New York Times from 2011-2017 and transformed them into a corpus to conduct a text analysis. How has the overall trend of public attention changed over the years, as seen by the number of articles relating to the ACA? What topics do people focus on when they discuss the ACA?

## Overview of the Affordable Care Act in the United States

The __Patient Protection and Affordable Care Act (ACA)__, also commonly known as Obamacare, was a major and wide-sweeping healthcare reform initiative enacted by Preseident Obama. The ACA was signed into law by Obama on March 23, 2010, and changes were gradually implemented over the next few years, though the majority of the provisions went into effect only on January 1, 2014.   

The major goal of the ACA is to expand health insurance and healthcare access to those who otherwise did not have access, including the 45 million Americans who were uninsured before the ACA took effect. In addition, it also covers wide-ranging regulations for the protection of healthcare consumer rights. Prior to the ACA, many low-income Americans we unable to afford any health insurance, while people with disabilities and pre-existing conditions may have been denied coverage by insurerers.   

__The key provisions of the ACA include__  

_Employer mandate_: Employers with more than 50 employees are required to provide health insurance to full-time workers, or would otherwise face a fine

_Individual mandate_: Individuals must obtain "minimum essential coverage" for healthcare, and those without health insurance must pay a fine in their annual tax returns to the IRS

_State and federal marketplaces_: These health insurance exchanges were set-up to enable people "shopping" for insurance to make easy comparisons among the insurance plans available, thus inreasing competition in the marketplace as well.

_Medicaid expansion_: The ACA expands medicaid to all individuals earning less than 138% of the federal poverty level (FPL), enabled first by federal funding and . After the Supreme Court ruled this requirement as unconstitutional (thus optional) in 2012, 31 states have elected to expand medicaid, while 19 have not.

_Federal subsidies_ to people earning between 138% and 400% of the FPL to afford monthly health insurance premiums.

_Regulations_ preventing insurance companies from discriminating in price and provision based on people's age, gender, or pre-existing conditions

These provisions of the ACA make health insurance accessible and affordable to people previously without coverage. In addition, people who already have health insurance would also benefit from increased choice, a cap on out-of-pocket expenses, as well as more comprehensive coverage, as insurance is now required to cover more basic treatments than before.

However, this increased coverage comes at a cost of increased premium growths for consumers, and increases in federal spending. The expansion of health insurance and healthcare coverage is projected to cost of the U.S. government up to $1.34 trillion over the next decade. It is thus no surprise that the States fall on various sides of the political spectrum when it comes to support for medicaid.

-----

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
rm(list=ls())

library(leaflet)
library(maps)
library(rgdal)
library(DT)
library(ggplot2)
library(ggthemes)
library(plotly)
library(magrittr)
library(readxl)
library(plyr)
library(dplyr)
library(tidyr)
library(readr)
library(stringi)
library(RColorBrewer)
library(countrycode)
require(gridExtra)
```


```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## IMPORT DATA

# General ACA information
aca_general_raw <- read_excel("Data/states_aca_general.xlsx", sheet=2, col_names=FALSE)

# States health outcomes
health_ind_raw <- read_excel("Data/states_health_ind.xlsx", sheet=4, col_names=FALSE)

# Insurance coverage data
  # Long format
insurance_long <- read_excel("Data/insurance-clean.xlsx", sheet=1, col_names=TRUE)
  # Wide format
insurance <- read_excel("Data/insurance-clean.xlsx", sheet=2, col_names=TRUE)

# Demographic information
demographics1 <- read.csv("Data/Population_Age_Income.csv", header = TRUE)
  # Provided with Assignment 4 (want the yearly population data)
demographics2 <- read_excel("Data/PopulationEstimates.xls")

# Election Results
election <- read_excel("Data/US_Presidential_Results_by_State_1828-2016.xlsx", sheet=2, col_names=FALSE)
```

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## CLEAN DATA

# List of states abbreviations
state_abbreviations <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY")


## CLEAN INSURANCE RATES DATA (WIDE)

# Drop unnecessary columns
insurance <- insurance[1:50, -1]

# Rename variables
colnames(insurance) <- c("state_abb", "state", "uninsured_num_2008", "uninsured_pct_2008", "uninsured_num_2009", "uninsured_pct_2009", "uninsured_num_2010", "uninsured_pct_2010", "uninsured_num_2011", "uninsured_pct_2011", "uninsured_num_2012", "uninsured_pct_2012", "uninsured_num_2013", "uninsured_pct_2013", "uninsured_num_2014", "uninsured_pct_2014", "uninsured_num_2015", "uninsured_pct_2015")

# Create variables for insurance rates
insurance$insured_pct_2008 <- 100 - insurance$uninsured_pct_2008
insurance$insured_pct_2009 <- 100 - insurance$uninsured_pct_2009
insurance$insured_pct_2010 <- 100 - insurance$uninsured_pct_2010
insurance$insured_pct_2011 <- 100 - insurance$uninsured_pct_2011
insurance$insured_pct_2012 <- 100 - insurance$uninsured_pct_2012
insurance$insured_pct_2013 <- 100 - insurance$uninsured_pct_2013
insurance$insured_pct_2014 <- 100 - insurance$uninsured_pct_2014
insurance$insured_pct_2015 <- 100 - insurance$uninsured_pct_2015

# Reorder variables
insurance <- insurance[ , c("state_abb", "state", "uninsured_num_2008", "uninsured_pct_2008", "insured_pct_2008", "uninsured_num_2009", "uninsured_pct_2009", "insured_pct_2009", "uninsured_num_2010", "uninsured_pct_2010", "insured_pct_2010", "uninsured_num_2011", "uninsured_pct_2011", "insured_pct_2011", "uninsured_num_2012", "uninsured_pct_2012", "insured_pct_2012", "uninsured_num_2013", "uninsured_pct_2013", "insured_pct_2013", "uninsured_num_2014", "uninsured_pct_2014", "insured_pct_2014", "uninsured_num_2015", "uninsured_pct_2015", "insured_pct_2015")]


## CLEAN INSURANCE RATES DATA (LONG)

# Rename variables
colnames(insurance_long) <- c("year", "state", "population", "uninsured_num", "uninsured_pct", "insured_num", "insured_pct")

# Add state abbreviations
  # First create a look-up table (based on wide data)
state_abb_lookup <- insurance[ , c("state", "state_abb")]
  # Then do the matching
insurance_long$state_abb <- state_abb_lookup$state_abb[match(insurance_long$state, state_abb_lookup$state)]
  # Reorder variable
insurance_long <- insurance_long[ , c(1:2, ncol(insurance_long), 3:(ncol(insurance_long)-1))]

# Order observations (by state, then by year)
insurance_long <- insurance_long[order(insurance_long$state, insurance_long$year), ]


## CLEAN ACA GENERAL DATA

# Keep only relevant rows and columns
aca_general <- aca_general_raw[6:57, 1:73]

# Label variables for ACA general data
names(aca_general) <- c("state", "unins_all_pct_10", "unins_all_pct_15", "unins_all_decr_pct", "unins_all_decr", "cov_emp", "cov_parents_plan", "lifetime_lim_preACA_tot", "lifetime_lim_preACA_child", "lifetime_lim_preACA_adultM", "lifetime_lim_preACA_adultF", "cov_private_tot", "cov_private_child", "cov_private_adultM", "cov_private_adultF", "premium_emp_avg_growth_pct_00to10", "premium_emp_avg_growth_pct_10to15", "premium_emp_savings_15", "premium_emp_savings_16", "MLR_rebate_beneficiaries_12", "MLR_rebate_amt_12", "MLR_rebate_beneficiaries_13", "MLR_rebate_amt_13", "MLR_rebate_beneficiaries_14", "MLR_rebate_amt_14", "MLR_rebate_beneficiaries_15", "MLR_rebate_amt_15", "MLR_rebate_amt_12to15", "medicaid_enroll_13", "medicaid_enroll_16", "medicaid_enroll_child_13", "medicaid_enroll_incr_13to16", "medicaid_full_duals", "medicaid_partial_duals", "medicaid_full_or_partial_duals", "state_has_expanded", "insurance_incr_medicaid", "cholest_scr_incr_medicaid", "mammogram_incr_medicaid", "papsmear_incr_medicaid", "clinic_care_incr_medicaid", "all_care_incr_medicaid", "phycisian_visit_ann_incr_medicaid", "depression_decr_medicaid", "good_health_incr_medicaid", "deaths_ann_decr_medicaid", "catastrophic_oop_ann_decr_medicaid", "indebted_ppl_decr_medicaid", "fed_spending_net_incr_inMil", "uncompensated_care_decr_inMil", "mental_substance_elig_medicaid_share", "mental_substance_elig_medicaid", "preexisting_condition_09", "cov_mkt_plan_16", "cov_mkt_16", "receive_tax_credit", "avg_tax_credit", "receive_cost_sharing", "avg_num_mkt_plans_avail_17", "cov_mkt_under75D_pct", "cov_mkt_under100D_pct", "cov_offmkt_elig_tax_credit", "rate_review_funds_to_state", "HIECP_grant_award_to_state", "medicare_enroll_16", "medicare_benef_donuthole", "medicare_benef_donuthole_savings_tot", "medicare_benef_donuthole_savings_avg", "medicare_partB_free_prev_services", "medicare_partB_free_prev_services_share", "medicare_incr_readmit_rate", "medicare_avoided_readmit", "accountable_care_org_num")

# Trim white space on state variable
aca_general$state <- stri_trim_both(aca_general$state)

# Change format of variables
  # Apart from T/F variable for "state has expanded"
aca_general$state_has_expanded[aca_general$state_has_expanded=="yes"] <- TRUE 
aca_general$state_has_expanded[aca_general$state_has_expanded=="no"] <- FALSE 
  # Convert others to numeric form
for (j in c(2:35, 37:73)) {
  aca_general[[j]] <- as.numeric(aca_general[[j]])
}

# Save USA data separately
aca_general_USA <- aca_general[aca_general$state=="United States", ]

# Drop USA and DC data
aca_general <- aca_general[!(aca_general$state %in% c("United States", "District of Columbia")), ]

# Add state abbreviations
aca_general$state_abb <- state_abbreviations
aca_general <- aca_general[ , c(1, ncol(aca_general), 2:(ncol(aca_general)-1))]

# Drop raw data
rm(aca_general_raw)


## CLEAN HEALTH INDICATOR DATA

# Keep only relevant rows and columns
health_ind <- health_ind_raw[4:6703, c(1, 3:6)]

# Label variables
names(health_ind) <- c("state", "measure", "year", "rate", "rank")

# Simplify date (years) - only keep last year if range was given
health_ind$year[health_ind$year=="07/2009 - 06/2012"] <- "2012"
health_ind$year[health_ind$year=="07/2010 - 06/2013"] <- "2013"
health_ind$year[health_ind$year=="07/2012 - 06/2015"] <- "2015"
health_ind$year[health_ind$year=="10/2013-9/2014"] <- "2014"
health_ind$year[health_ind$year=="2008-09"] <- "2009"
health_ind$year[health_ind$year=="2010-11"] <- "2011"
health_ind$year[health_ind$year=="2011-12"] <- "2012"
health_ind$year[health_ind$year=="2011/12"] <- "2012"
health_ind$year[health_ind$year=="2012-13"] <- "2013"
health_ind$year[health_ind$year=="2013-14"] <- "2014"
health_ind$year[health_ind$year=="2013(Q2-Q4)"] <- "2013"
health_ind$year[health_ind$year=="2014-15"] <- "2015"
health_ind$year[health_ind$year=="2015(Q2-Q4)"] <- "2015"

# Export to excel to manually label variable names
unique_measures <- unique(as.factor(health_ind$measure))
  # Copied in from excel sheet
unique_measures_relabelled <- c("a.summary_access", "a.unins_adult", "a.unins_child", "a.no_care_bc_cost_adult", "a.high_OOP_relative_under65", "a.at_risk_no_routine_doc_adult", "a.no_dental_adult", "q.summary_prev_treat", "q.with_usual_care_adult", "q.with_cancer_screening_adult", "q.with_vaccines_adult", "q.with_medical_home_child", "q.with_prev_medical_dental_child", "q.with_mental_healthcare_child", "q.with_vaccines_infant", "q.drug_should_avoid_medicare", "q.drug_should_avoid_3conditions_medicare", "q.good_health_provider_medicare", "q.mortality_4conditions_medicare", "q.CLABSI_infection_ratio", "q.info_recovery_hospitalized", "q.good_hospital_staff_hospitalized", "q.improve_mobility_homehealth", "q.improved_wounds_homehealth", "q.sores_NHres", "q.antipsychotic_med_NHres", "u.summary_avoidable_hosp_cost", "u.hosp_asthma_child", "u.hosp_ambulatory_65to74yrs", "u.hosp_ambulatory_above75yrs", "u.30day_hosp_readmit_medicare", "u.30day_hosp_readmit_NHres", "u.hosp_6mos_NHres", "u.hosp_medicare_homehealth", "u.avoidable_ER_medicare", "u.tot_reimb_employer_ins", "u.tot_reimb_medicare", "h.summary_healthy_lives", "h.deaths_amenable", "h.yrs_lost_potential_life_before75", "h.deaths_breast_cancer_F", "h.deaths_colorectal_cancer", "h.deaths_suicide", "h.deaths_infant_mortality", "h.poor_health_adult", "h.smoke_adult", "h.obese_adult", "h.obese_child", "h.poor_dental_adult", "u.premium_emp_private", "u.premium_emp_private_unadj", "u.reimb_medicare_unadj", "u.deaths_amenable_black", "u.deaths_amenable_white", NA, "q.with_prev_screening_above50yrs", NA)

# Replace variable names with abbreviated version
for (i in 1:57) {
  health_ind$measure <- replace(health_ind$measure, health_ind$measure==unique_measures[i], unique_measures_relabelled[i])
}

# Remove the NA's
health_ind <- health_ind[!is.na(health_ind$measure), ]

# Remove duplicates
  # By state, measure.year
health_ind <- health_ind[!duplicated(health_ind[ , 1:3]), ]

# Interact measure and year
health_ind$measure_year <- interaction(health_ind$measure, health_ind$year)

# Convert to character variable
health_ind$measure_year <- as.character(health_ind$measure_year)

# Pre-reshape: Store existing data as "long"
health_ind_long <- health_ind

# Reshape wide
health_ind <- as.data.frame(cbind(health_ind$state, health_ind$measure_year, health_ind$rate))
  # Drop rank data for now
names(health_ind) <- c("state", "measure_year", "rate")
health_ind <- reshape(health_ind,
                        idvar = "state",
                        timevar = "measure_year",
                        direction = "wide")

# Drop "rate" from the name
names(health_ind)[2:ncol(health_ind)] <- substr(names(health_ind), 6, nchar(names(health_ind)))[2:ncol(health_ind)]

# Drop variables who have all entries as NA
  # Due to 'unbalanced panel' in reshaping
  # First create a vector to record if variable is all NA
variable_all_NA <- vector(mode="logical", length=ncol(health_ind))
for (j in 1:ncol(health_ind)) {
  if(sum(is.na(health_ind[[j]]))==52) {
    variable_all_NA[j] <- TRUE
  }
}
  # Only keep the variables that are NOT all NA
health_ind <- health_ind[ , !variable_all_NA]

# Change format of variables
  # To character variable
health_ind$state <- as.character(health_ind$state)
  # To numeric variable
for (j in c(2:164)) {
  health_ind[[j]] <- as.character(health_ind[[j]])
  health_ind[[j]] <- as.numeric(health_ind[[j]])
}

# Save USA data separately
health_ind_USA <- health_ind[health_ind$state=="United States", ]

# Drop USA and DC data
health_ind <- health_ind[!(health_ind$state %in% c("United States", "District of Columbia")), ]

# Label state abbreviations
health_ind$state_abb <- state_abbreviations
health_ind <- health_ind[ , c(1, ncol(health_ind), 2:(ncol(health_ind)-1))]

# Drop raw data
rm(health_ind_raw)


## CLEAN DEMOGRAPHIC (1) DATA

# Drop unnecessary row and column
demographics1 <- demographics1[-1,-1]

# Rename variables
colnames(demographics1) <- c("state","income","population", "ppl_age0to4","ppl_age5to9","ppl_age10to14","ppl_age15to19","ppl_age20to24", "ppl_age25to29","ppl_age30to34","ppl_age35to39","ppl_age40to44", "ppl_age45to49","ppl_age50to54","ppl_age55to59","ppl_age60to64", "ppl_age65to69","ppl_age70to74","ppl_age75to79","ppl_age80to84","ppl_age85plus")

# Drop observations DC and PR
demographics1 <- demographics1[!demographics1$state=="District of Columbia" & !demographics1$state=="Puerto Rico", ]

# Label state abbreviations
demographics1$state_abb <- state_abbreviations
demographics1 <- demographics1[ , c(1, ncol(demographics1), 2:(ncol(demographics1)-1))]

# Change format of variables
  # To character variable
demographics1$state <- as.character(demographics1$state)
  # To numeric variable
for (j in c(3:ncol(demographics1))) {
  demographics1[[j]] <- as.character(demographics1[[j]])
  demographics1[[j]] <- as.numeric(demographics1[[j]])
}

# BEA regions
demographics1$BEA_region <- ""
demographics1$BEA_region[demographics1$state_abb %in% c("ME", "NH", "VT", "MA", "CT", "RI")] <- "New England"
demographics1$BEA_region[demographics1$state_abb %in% c("NY", "NJ", "PA", "MD", "DE")] <- "Mideast"
demographics1$BEA_region[demographics1$state_abb %in% c("WI", "IL", "MI", "IN", "OH")] <- "Great Lakes"
demographics1$BEA_region[demographics1$state_abb %in% c("WV", "KY", "VA", "TN", "NC", "SC", "AR", "LA", "MS", "AL", "GA", "FL")] <- "Southeast"
demographics1$BEA_region[demographics1$state_abb %in% c("ND", "SD", "NE", "KS", "MN", "IA", "MO")] <- "Plains"
demographics1$BEA_region[demographics1$state_abb %in% c("MT", "ID", "WY", "UT", "CO")] <- "Rocky Mountains"
demographics1$BEA_region[demographics1$state_abb %in% c("AZ", "NM", "TX", "OK")] <- "Southwest"
demographics1$BEA_region[demographics1$state_abb %in% c("WA", "OR", "CA", "NV", "AK", "HI")] <- "Far West"

# Tag income level (category)
demographics1$income_level <- ""
demographics1$income_level[demographics1$income >= 60000] <- "High"
demographics1$income_level[demographics1$income >= 50000 & demographics1$income < 60000] <- "Upper Middle"
demographics1$income_level[demographics1$income >= 45000 & demographics1$income < 50000] <- "Lower Middle"
demographics1$income_level[demographics1$income < 45000] <- "Low"
  # Convert to factor variable
demographics1$income_level <- factor(demographics1$income_level, levels = c("Low", "Lower Middle", "Upper Middle", "High"))


## CLEAN DEMOGRAPHIC (2) DATA

colnames(demographics2) <- demographics2[2, ]

# Drop first few rows (non-observations)
demographics2 <- demographics2[-(1:2), ]

# Drop county-level data,  keep state-level data only
demographics2 <- demographics2[demographics2$Area_Name == "United States" | demographics2$State != lag(demographics2$State), ]

# Keep only columns of population from 2010 to 2015
demographics2 <- demographics2[ , c("State", "Area_Name", "POP_ESTIMATE_2010", "POP_ESTIMATE_2011", "POP_ESTIMATE_2012", "POP_ESTIMATE_2013", "POP_ESTIMATE_2014", "POP_ESTIMATE_2015")]

# Change format of variables
  # To character variable
for (j in c(3:8)) {
  demographics2[[j]] <- as.numeric(demographics2[[j]])
}

# Rename variables
colnames(demographics2) <- c("state_abb", "state", "population_2010", "population_2011", "population_2012", "population_2013", "population_2014", "population_2015")


## CLEAN ELECTION DATA

# Keep only 2012 and 2008 data, and data for the 50 states
election <- election[ 4:53 , c("X0", "X5", "X6", "X9", "X10")]

# Rename variables
colnames(election) <- c("state", "Dem_pct_2012", "Rep_pct_2012", "Dem_pct_2008", "Rep_pct_2008")

# Add state abbreviations
  # Match with look-up table
election$state_abb <- state_abb_lookup$state_abb[match(election$state, state_abb_lookup$state)]

# Change format of variables
  # To numeric variable
for (j in c(4:6)) {
  election[[j]] <- as.numeric(election[[j]])
}

# Calculate winning party for each state
election$party_2012 <- ifelse(election$Dem_pct_2012 > election$Rep_pct_2012, "Democratic", "Republican")
election$party_2008 <- ifelse(election$Dem_pct_2008 > election$Rep_pct_2008, "Democratic", "Republican")

# Reorder variables
election <- election[ , c("state", "state_abb", "Dem_pct_2012", "Rep_pct_2012", "party_2012", "Dem_pct_2008", "Rep_pct_2008", "party_2008")]
```


# (1) Health Insurance Coverage

__Figure 1__

Access our Shiny App on [Uninsured Rates in the United States from 2008-2015](https://gracekongyx.shinyapps.io/health_insurance_shiny/).
   
   

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## SOME DATA PREPARATION

# Look-up information into insurance and insurance_long, on:
  # Party (2012)
insurance_long$party_2012 <- election$party_2012[match(insurance_long$state, election$state)]
insurance$party_2012 <- election$party_2012[match(insurance$state, election$state)]
  # BEA region
insurance_long$BEA_region <- demographics1$BEA_region[match(insurance_long$state, demographics1$state)]
insurance$BEA_region <- demographics1$BEA_region[match(insurance$state, demographics1$state)]
  # Income level
insurance_long$income_level <- demographics1$income_level[match(insurance_long$state, demographics1$state)]
insurance$income_level <- demographics1$income_level[match(insurance$state, demographics1$state)]

# Also find nation-wide trend for insurance
insurance_long_US <- data.frame("year" = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015), "uninsured_pct" = c(16.8, 17.5, 18.2, 17.2, 16.9, 16.6, 13.3, 10.5))
```

__Figure 2__
   
The following figure displays the trend in the uninsured rate between 2008 and 2015. Results are shown for the USA average (in black), as well as for every individual state (in color). 

We see a clear downward trend in the uninsured rate once most of the Affordable Care Act (ACA) provisions were implemented in 2014. The overall US uninsured rate decreased from 18.2% in 2010 to 10.5% in 2015.

However, note that the disparities in uninsured rates largely continue to persist among the states, with regional patterns (using Bureau of Economic Analysis (BEA) regions). States of the Southeast and Southwest consistently had the highest uninsured rates, while states of New England consistenely had the lowest insured or highest health insurance coverage rates.
   
```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 2

plot1 <- (ggplot(insurance_long, aes(year, uninsured_pct)) 
          + geom_line(aes(color = BEA_region, group = state, label = uninsured_num))
          + scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015))
          + scale_y_continuous(breaks = c(5, 10, 15, 20, 25, 30))
          + labs(x = "Year", y = "% Uninsured", color = "Region")
          + geom_line(data = insurance_long_US, lwd = 0.8)
          + geom_vline(xintercept = 2010, linetype = 2, color = "gray30")
          + geom_vline(xintercept = 2014, linetype = 2, color = "gray30")
          + annotate("text",x = 2009.4, y = 25, label = "Obama \nsigns ACA", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2013.4, y = 24.5, label = "Most ACA \nprovisions \ntake effect", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2014.2, y = 11, label = "USA \nAverage", color = "black",fontface = 2, size = 3)
          + ggtitle("Change in Percentage Uninsured by State, \nfrom 2008 to 2015")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0, size = 12)))
ggplotly(plot1, dynamicTicks = FALSE, tooltip = c("state", "year", "uninsured_pct", "uninsured_num"))
```
   
__Figure 3__

In the following plot, I explore the difference between Democratic and Republican states, based on the 2012 US Presidential Election results. Generally, the democratic states consistently have had largely have a lower uninsured rate than the Republican states, though both groups experienced declines in the uninsured rates in 2014. The rate of decline of uninsured rates in 2014 for many Republican states appeared to exceed that of the Democratic states, mainly because of their higher starting point.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 3
plot2 <- (ggplot(insurance_long, aes(year, uninsured_pct)) 
          + geom_line(aes(color = party_2012, group = state, label = uninsured_num))
          + scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015))
          + scale_y_continuous(breaks = c(5, 10, 15, 20, 25, 30))
          + scale_color_manual(values = c("steel blue", "firebrick1")) 
          + labs(x = "Year", y = "% Uninsured", color = "Party \n(2012 Winner)")
          + geom_line(data = insurance_long_US, lwd = 1)
          + geom_vline(xintercept = 2010, linetype = 2, color = "gray30")
          + geom_vline(xintercept = 2014, linetype = 2, color = "gray30")
          + annotate("text",x = 2009.4, y = 25, label = "Obama \nsigns ACA", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2013.4, y = 24.5, label = "Most ACA \nprovisions \ntake effect", color = "black",fontface = 2, size = 3)
          + annotate("text",x = 2014.2, y = 11, label = "USA \nAverage", color = "black",fontface = 2, size = 3)
          + ggtitle("Change in Percentage Uninsured by Party, \nfrom 2008 to 2015")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0, size = 12)))
ggplotly(plot2, dynamicTicks = FALSE,  tooltip = c("state", "year", "uninsured_pct", "uninsured_num"))
```

__Figure 4__

This plot shows insured rates in 2015 at a glance, looking at how they differ by party (2012 winner), BEA region and income level of the states. 

In terms of correlations, it appears that being a Democratic state is a strong predictor of a higher insurance rate. States in New England, the Mideast and the Great Lakes perform well in insurance rates, while states in the Southwest, Rocky Mountains and Southeast perform the poorest. As expected, high-income states have the highest insurance rates as more people would be able to afford insurance. Lower and lower-middle income states have lower coverage rates, but it is infact the lower-middle income states that have the lowest insurance rates.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 4

# Subset relevant variables
insurance_2015 <- insurance[ , c("state", "uninsured_pct_2015", "insured_pct_2015", "party_2012", "BEA_region", "income_level")]

# Boxplots by party
plot3a <- (ggplot(insurance_2015, aes(x = reorder(party_2012, insured_pct_2015), y = insured_pct_2015)) 
          + geom_boxplot(aes(fill = party_2012, label = insured_pct_2015), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = c("steel blue", "firebrick"))
          + ggtitle("Democratic vs. Republican States")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.37, size = 12), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_blank()))

# Boxplots by BEA region
plot3b <- (ggplot(insurance_2015, aes(x = reorder(BEA_region, insured_pct_2015), y = insured_pct_2015)) 
          + geom_boxplot(aes(fill = BEA_region, label = insured_pct_2015), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = brewer.pal(n = 8, name = "Dark2"))
          + ggtitle("By Bureau of Economic Analysis (BEA) Region")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.24, size = 12), legend.position = "none", axis.title.x = element_blank(), axis.title.y = element_blank()))

# Boxplots by income level
plot3c <- (ggplot(insurance_2015, aes(x = income_level, y = insured_pct_2015)) 
          + geom_boxplot(aes(fill = income_level, label = insured_pct_2015), color = "gray30")
          + coord_flip()
          + scale_fill_manual(values = brewer.pal(n = 4, name = "Dark2"))
          + labs(x = "", y = "% Insured (2015)")
          + ggtitle("By State's Income Level")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.39, size = 12), legend.position = "none", axis.title.y = element_blank()))

# Combine plots vertically
grid.arrange(plot3a, plot3b, plot3c, nrow = 3, top = "2015 Insurance Rates at a Glance")
```



```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## SOME DATA RESHAPING

# Reshape aca_general (long), keeping the interesting quantities
  # Import uninsured number (2015)
aca_general$unins_all_15 <- insurance$uninsured_num_2015[match(aca_general$state, insurance$state)]
  # Keep only select variables
aca_general_long <- aca_general[ , c("state", "state_abb", "cov_emp", "cov_mkt_plan_16", "medicaid_enroll_16", "medicare_enroll_16", "unins_all_15")]
  # The actual reshaping
aca_general_long <- reshape(aca_general_long, 
                      varying = c("cov_emp", "cov_mkt_plan_16", "medicaid_enroll_16", "medicare_enroll_16", "unins_all_15"),
                      v.names = "Number",
                      timevar = "Source",
                      times = c("Employer", "Marketplace", "Medicaid / CHIP", "Medicare", "Uninsured"),
                      new.row.names = 1:1000,
                      direction = "long")

# Look-up information into aca_general_long, on:
  # Party (2012)
aca_general_long$party_2012 <- election$party_2012[match(aca_general_long$state, election$state)]
  # BEA region
aca_general_long$BEA_region <- demographics1$BEA_region[match(aca_general_long$state, demographics1$state)]
  # Income level
aca_general_long$income_level <- demographics1$income_level[match(aca_general_long$state, demographics1$state)]
  # Population
aca_general_long$population_2015 <- demographics2$population_2015[match(aca_general_long$state, demographics2$state)]

# Calculate percentage from number and population
aca_general_long$percentage <- aca_general_long$Number / aca_general_long$population_2015 * 100
```
  
  
__Figure 5__

Next, we investigate the main health insurance and coverage sources, namely employer coverage, the individual marketplace, Medicaid / Childrens' Health Insurance Program (CHIP), and Medicare for the aged. We broke down the results by party and income level, taking a simple (unweighted) average across the states in the categories. We observe that employer coverage rates increase as income level increases, especially for Democratic states. Democratic and low income states have more people on Medicaid / CHIP than Republican and low income states.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 5

aca_general_long_by_party_income <- ddply(aca_general_long, c("party_2012", "income_level", "Source"), summarise, avg_pct = mean(percentage))

plot4 <- (ggplot(aca_general_long_by_party_income, aes(income_level, avg_pct, fill = Source))
          + geom_col(position = "dodge") + facet_grid(party_2012 ~ .)
          + labs(x = "Income Level", y = "Average % in State", color = "Source of \nInsurance")
          + scale_fill_manual(values = brewer.pal(n = 6, name = "Dark2")[c(1:3,6,4)])
          + ggtitle("Sources of Health Insurance, \nby Party and Income Level")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0, size = 12)))
ggplotly(plot4, tooltip = c("income_level", "Source", "avg_pct"))
```


```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PREPARATION FOR FURTHER PLOTS (INCLUDING DATA TABLE)

# Want to combine information on insurance coverage with key information on health outcomes/access and affordability

key_indicators <- state_abb_lookup

## Merge in information to the data table on the following

# Demographic Information
  # Income level
key_indicators$income_level <- demographics1$income_level[match(key_indicators$state_abb, demographics1$state_abb)]

# Insurance Coverage
  # Percentage insured (2015)
key_indicators$insured_pct_2015 <- insurance$insured_pct_2015[match(key_indicators$state_abb, insurance$state_abb)]
  # Percentage insured (2014)
key_indicators$insured_pct_2014 <- insurance$insured_pct_2014[match(key_indicators$state_abb, insurance$state_abb)]
  # Indicator: Whether state has expanded medicare
key_indicators$state_has_expanded <- aca_general$state_has_expanded[match(key_indicators$state_abb, aca_general$state_abb)]

# Health outcomes and access
  # Mortality amenable to health care, deaths per 10000 population
key_indicators$deaths_amenable_2014 <- health_ind$h.deaths_amenable.2014 [match(key_indicators$state_abb, health_ind$state_abb)]
  # Years of Potential Life Lost before 75
key_indicators$years_life_lost <- health_ind$h.yrs_lost_potential_life_before75.2014 [match(key_indicators$state_abb, health_ind$state_abb)]
  # Adults with a usual source of care (%)
key_indicators$usual_care_2015 <- health_ind$q.with_usual_care_adult.2015 [match(key_indicators$state_abb, health_ind$state_abb)]
  # Adults with age/gender-appropriate cancer screenings (%)
key_indicators$with_cancer_screening <- health_ind$q.with_cancer_screening_adult.2014 [match(key_indicators$state_abb, health_ind$state_abb)]
  # Adults with age-appropriate vaccines (%)
key_indicators$with_vaccines <- health_ind$q.with_vaccines_adult.2015 [match(key_indicators$state_abb, health_ind$state_abb)]

# Affordability and cost-efficiency
  # Individuals under age 65 with high OOP medical costs relative to annual household income
key_indicators$high_OOP_relative <- health_ind$a.high_OOP_relative_under65.2015 [match(key_indicators$state_abb, health_ind$state_abb)]
  # Average annual growth in family premiums for employer coverage (between 2010 and 2015)
key_indicators$premium_ann_growth_10_15 <- aca_general$premium_emp_avg_growth_pct_10to15[match(key_indicators$state_abb, aca_general$state_abb)]
  # Marketplace consumers who could select a plan for less than $100
key_indicators$IM_plan_under_100 <- aca_general$cov_mkt_under100D_pct[match(key_indicators$state_abb, aca_general$state_abb)]
  # Net increase in federal spending (millions)
key_indicators$incr_fed_spending_mil <- aca_general$fed_spending_net_incr_inMil[match(key_indicators$state_abb, aca_general$state_abb)]
```

  
# (2) Health Outcomes and Access

__Figure 6__

Next, we are interested to find out how health insurance rates correlates with health outcomes. In particular, we can look at the measure of mortality amenable to healthcare intervention - the number of deaths that could have been prevented by healthcare in the year 2014. We plot avoidable mortality (2014) against insurance rates (2014) and observe a negative relationship as expected. States that have higher rates of insurance also have better health in this respect. However, we also control for income levels as low income levels would be a major predictor of poor health, as confirmed in our plot. Yet, within each income level, the negative relationship remains, and it is especially pronounced among the upper-middle income countries. 

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 6
plot5 <- (ggplot(key_indicators, aes(insured_pct_2014, deaths_amenable_2014))
          + geom_point(aes(color = income_level, shape = state_has_expanded), size = 2)
          + geom_smooth(aes(color = income_level, group = income_level), method = "lm", lwd = 1, se = FALSE)
          + geom_smooth(color = "black", method = "lm", linetype = 2, lwd = 1, se = FALSE)
          + labs(x = "% Insured", y = "Deaths Preventable by Healthcare Intervention \n(per 100,000 People)", color = "Income \nLevel", shape = "State has \nExpanded \nMedicaid")
          + scale_color_manual(values = brewer.pal(n = 4, name = "Dark2"))
          + ggtitle("Relationship Between Avoidable Mortality and \nHealth Insurance Coverage Among States")
          + theme_minimal()
          + theme(plot.title = element_text(face="bold", hjust = 0.5, size = 12)))
plot5
```


```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## DATA PROCESSING FOR MAPS

# Obtain shape files of US states
map_states = map("state", fill = TRUE, plot = FALSE)

# Standardize name format with rest of assignment
  # Function to conver to proper case
properCase <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
  # Convert to proper case
map_states$state <- sapply(map_states$names, properCase)
  # Rename the main component of states with multiple parts in the map
map_states$state[map_states$state == "Massachusetts:main"] <- "Massachusetts"
map_states$state[map_states$state == "Michigan:south"] <- "Michigan"
map_states$state[map_states$state == "New York:main"] <- "New York"
map_states$state[map_states$state == "North Carolina:main"] <- "North Carolina"
map_states$state[map_states$state == "Virginia:main"] <- "Virginia"
map_states$state[map_states$state == "Washington:main"] <- "Washington"
  # Match with state abbreviations
map_states$state_abb <- state_abb_lookup$state_abb[match(map_states$state, state_abb_lookup$state)]
  # Then strip off all the extra location information for states with multiple parts (leaving just the state name)
for (i in 1:length(map_states$state)) {
  map_states$state[i] <- unlist(strsplit(map_states$state[i], ":"))[1]
}

# Get coordinates of state centers
states_centers <- state.center
state.center <- cbind(states_centers, state_abb_lookup)

# Health outcomes: Import key indicators data, matched by state
    # Remember to later match information based on state, not state abbreviation (due to how we labelled above)
map_states$deaths_amenable_2014 <- key_indicators$deaths_amenable_2014[match(map_states$state, key_indicators$state)]
map_states$usual_care_2015 <- key_indicators$usual_care_2015[match(map_states$state, key_indicators$state)]
map_states$with_cancer_screening <- key_indicators$with_cancer_screening[match(map_states$state, key_indicators$state)]
map_states$with_vaccines <- key_indicators$with_vaccines[match(map_states$state, key_indicators$state)]

# Cost outcomes: Import key indicators data, matched by state
    # Remember to later match information based on state, not state abbreviation (due to how we labelled above)
map_states$high_OOP_relative <- key_indicators$high_OOP_relative[match(map_states$state, key_indicators$state)]
map_states$premium_ann_growth_10_15 <- key_indicators$premium_ann_growth_10_15[match(map_states$state, key_indicators$state)]
map_states$IM_plan_under_100 <- key_indicators$IM_plan_under_100[match(map_states$state, key_indicators$state)]
map_states$incr_fed_spending_mil <- key_indicators$incr_fed_spending_mil[match(map_states$state, key_indicators$state)]
```


__Figure 7__

The following maps how the states performed in various health outcomes, namely preventable mortality, access to a usual source of care, rate of age-appropriate cancer screenings in adults, and rate of age-appropriate vaccinations in adults. A green color incidates a comparatively more favorable outcome, while a red color indicates a less favorable outcome.

We note that the Northeast (New England) states perform consistently well across all health outcomes, while the Southeast states, as well as Texes, peform consistently poorly on all except for vaccination rates. The Western and Rocky Mountains states have a low avoidable mortality rate, despite having low rates of access to usual care and of cancer screening and vaccinations. In the measure of adult vaccination rates, the trend of the South faring poorer on health is reversed, as they do comparatively well in vaccinations.
  

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 7: LEAFLET MAP - HEALTH OUTCOMES

(leaflet(map_states) %>% 
    setView(lat=39.8282, lng=-96 , zoom=4) %>%
    addPolygons(color = "#333333", weight = 1, smoothFactor = 0.5, fillOpacity = 0) %>%
    addPolygons(group = "Preventable Deaths", fillColor = ~colorQuantile("RdYlGn", -deaths_amenable_2014)(-deaths_amenable_2014), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>Preventable Mortality</b>", "<br/>", "<b>per 100,000: </b>", map_states$deaths_amenable_2014)) %>%
    addPolygons(group = "Access to Usual Care", fillColor = ~colorQuantile("RdYlGn", usual_care_2015)(usual_care_2015), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>Access to Usual</b>", "<br/>", "<b>Source of Care: </b>", map_states$usual_care_2015, "%")) %>%
    addPolygons(group = "Cancer Screenings Rate", fillColor = ~colorQuantile("RdYlGn", with_cancer_screening)(with_cancer_screening), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>Cancer Screenings</b>", "<br/>", "<b>Rate: </b>", map_states$with_cancer_screening, "%")) %>%
    addPolygons(group = "Adult Vaccination Rate", fillColor = ~colorQuantile("RdYlGn", with_vaccines)(with_vaccines), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>Vaccination</b>", "<br/>", "<b>Rate: </b>", map_states$with_vaccines, "%")) %>%
    addLabelOnlyMarkers(data = filter(state.center, state_abb!="AK" & state_abb!="HI"), lng = ~x, lat = ~y, label = ~state_abb, labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T)) %>%
    addLayersControl(
      baseGroups = c("Preventable Deaths", "Access to Usual Care", "Cancer Screenings Rate", "Adult Vaccination Rate"),
      options = layersControlOptions(collapsed = FALSE))) 
```


# (3) Affordability and Cost Outcomes

__Figure 8__

Similarly, we map how the states performed in various financial and cost outcomes, such as percentage of people with high OOP costs relative to household income, premium growth rates, affordability of the marketplace plans, and increases in federal spending related to the state. 

We notice that states that do better in containing OOP costs tend to do worse in containing premium and marketplace plan rates as well as federal spending. These include some states like New York, New Jersey and Ohio in the Northeast and Midwest (Great Lakes). A notable exception is Massachussets, which performs well across the financial measures, possibly because of the headstart it got with "Romney-care", which actually served as a model for Obamacare.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 8: LEAFLET MAP - AFFORDABILITY / COST OUTCOMES

(leaflet(map_states) %>% 
    setView(lat=39.8282, lng=-96 , zoom=4) %>%
    addPolygons(color = "#333333", weight = 1, smoothFactor = 0.5, fillOpacity = 0) %>%
    addPolygons(group = "Contained OOP Costs", fillColor = ~colorQuantile("RdYlGn", -high_OOP_relative)(-high_OOP_relative), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>High OOP Relative</b>", "<br/>", "<b>to Income: </b>", map_states$high_OOP_relative, "%")) %>%
    addPolygons(group = "Premium Growth Rate", fillColor = ~colorQuantile("RdYlGn", -premium_ann_growth_10_15)(-premium_ann_growth_10_15), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>Annual Premium</b>", "<br/>", "<b>Growth Rate: </b>", map_states$premium_ann_growth_10_15, "%")) %>%
    addPolygons(group = "Affordable Individual Marketplace Plan", fillColor = ~colorQuantile("RdYlGn", IM_plan_under_100)(IM_plan_under_100), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>Marketplace Consumers</b>", "<br/>", "<b>Who Can Select Plan</b>", "<br/>","<b>Under $100: </b>", map_states$IM_plan_under_100 * 100, "%")) %>%
    addPolygons(group = "Increase in Federal Spending", fillColor = ~colorQuantile("RdYlGn", -incr_fed_spending_mil)(-incr_fed_spending_mil), smoothFactor = 0.5, stroke = FALSE, fillOpacity = 0.6, popup = paste("<b>State: </b>", map_states$state, "<br/>", "<b>Net Increase in Federal</b>", "<br/>", "<b>Spending</b>: $", map_states$incr_fed_spending_mil, "million")) %>%
    addLabelOnlyMarkers(data = filter(state.center, state_abb!="AK" & state_abb!="HI"), lng = ~x, lat = ~y, label = ~state_abb, labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T)) %>%
    addLayersControl(
      baseGroups = c("Contained OOP Costs", "Premium Growth Rate", "Affordable Individual Marketplace Plan", "Increase in Federal Spending"),
      options = layersControlOptions(collapsed = FALSE))) 
```


# Summary of States' Performance

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
# Create data table reporting values
data_table_values <- key_indicators[ , c("state", "insured_pct_2015", "deaths_amenable_2014", "usual_care_2015", "high_OOP_relative", "premium_ann_growth_10_15")]

# Create data table reporting rank
data_table_rank <- data_table_values

# Convert each function to rank
  # Rank 1 is always more favorable (regardless of definition of variable)    
data_table_rank$r_insured_pct_2015 <- rank(-data_table_values$insured_pct_2015, na.last = "keep", ties.method = "min")
data_table_rank$r_deaths_amenable_2014 <- rank(data_table_values$deaths_amenable_2014, na.last = "keep", ties.method = "min")
data_table_rank$r_usual_care_2015 <- rank(-data_table_values$usual_care_2015, na.last = "keep", ties.method = "min")
data_table_rank$r_high_OOP_relative <- rank(data_table_values$high_OOP_relative, na.last = "keep", ties.method = "min")
data_table_rank$r_premium_ann_growth_10_15 <- rank(data_table_values$premium_ann_growth_10_15, na.last = "keep", ties.method = "min")
  # Remove value variables
data_table_rank$insured_pct_2015 <- NULL
data_table_rank$deaths_amenable_2014 <- NULL
data_table_rank$usual_care_2015 <- NULL
data_table_rank$high_OOP_relative <- NULL
data_table_rank$premium_ann_growth_10_15 <- NULL
```

The following data table presents the rankings of the states on the following insurance coverage, health and affordability outcomes, providing a summary of the states' performance:

1) Percentage of population insured

2) Deaths preventable by healthcare intervention (per 100,000 people)

3) Proportion of adults with access to a usual source of care

4) Proportion of people with high out-of-pocket costs (relative to household income)

5) Annual rate of health insurance premium growth from 2010 to 2015

A rank of 1 always indicates the more favorable outcome (i.e. higher insurance and treatment access rate, lower preventable mortality, OOP costs and premium growth)

__Figure 9__

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
## PLOT 9: DATA TABLE RANKING THE STATES ON VARIOUS MEASURES
datatable(data_table_rank, rownames = FALSE, colnames = c("State", "Percentage Insured", "Fewer Preventable Deaths", "Usual Care Access", "OOP Contained", "Lower Annual Premium Growth"), caption = "Rank of the 50 States in Insurance Coverage and Health Outcomes")
```

# (4) Perceptions of ACA (Text Analysis)

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
library(tm)
library(tm.plugin.lexisnexis)
library(readxl)
library(gtools)  # for smartbind
library(dplyr)   # for data_frame
library(lubridate)   # for date formatting
library(stringr)  
library(tools)  # Title case
library(quanteda)
library(ggplot2)
library(quanteda)
library(stringr)
library(tm)
library(qdap)
library(SnowballC)
library(dplyr)
library(tidytext)
library(wordcloud)

# Combine CSV and HTML Files
data <- read_excel("LexisNexis/NYTimes_Metadata.xlsx")
colnames(data) <- tolower(colnames(data))

# Correct data
data$date <- substr(parse_date_time(data$date, c("mdy")),1,10)
data$author <- toTitleCase(tolower(data$byline))
data$byline <- NULL

## Get Text files
source1 <- LexisNexisSource("LexisNexis/The_New_York1.html")
source2 <- LexisNexisSource("LexisNexis/The_New_York2.html")

corpus1 <- Corpus(source1, readerControl = list(language = NA))
corpus2 <- Corpus(source2, readerControl = list(language = NA))

corpus <- c(corpus1, corpus2)

# Convert to quanteda corpus
corpus <- quanteda::corpus(corpus)

## Add Metadata
# Check: match(data$headline, corpus$documents$heading)
corpus$documents$datetimestamp <- substring(corpus$documents$datetimestamp, 1,4)
corpus$documents$date <- corpus$documents$datetimestamp
corpus$documents$description <- corpus$documents$id

# options(width = 200)
# kwic(corpus, "Trump")
# 
save(corpus, file="nytimes.rda")
```

__Figure 10__

We could see that people discussed ACA a lot when Obama first signed it, and the election in 2017 made it a hot topic again. And not suprisingly, people always talk about Trump when they talk about ACA.

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
dat <- data.frame(date =corpus$documents$datetimestamp, levels=c(1:1000))
ggplot(data=dat, aes(x=date))+geom_bar(stat="count")+ggtitle("Number of Articles Mentioning ACA by Year")+theme_tufte()+ylab("Number of articles") + xlab("year") + theme(plot.title = element_text(face="bold", hjust = 0.5, size = 12))

```

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
load("nytimes.rda")
dfmtotal <- dfm(corpus, remove = stopwords("english"), stem = TRUE, removePunct = TRUE, removeNumbers = TRUE, tolower = TRUE, verbose = TRUE)
dfmtotal[, 1:5]
head(stopwords("english"), 20)
freq<-topfeatures(dfmtotal, 20)
wf <- data.frame(word=names(freq), freq=freq)   
```

__Figure 11__

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
p <- ggplot(subset(wf, freq>50), aes(word, freq))    
p <- p + geom_bar(stat="identity")
p <- p + geom_text(aes(label=freq), vjust=-0.2)+ scale_x_discrete(limits= wf$word)   
p <- p + theme(axis.text.x=element_text(angle=45, hjust=1), plot.title = element_text(face="bold", hjust = 0.5, size = 12))+ggtitle("Word Frequency from Articles on ACA, 2011-2017")+theme_tufte()+ylab("Frequency of Stem") + xlab("Top 30 Used Stem")
p
```

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
load('nytimes.rda')
NYT_source <- VectorSource(corpus$documents[,1])
NYT <- VCorpus(NYT_source)
documents <- corpus$documents
removeNumPunct <- function(x){gsub("[^[:alpha:][:space:]]*", "", x)}
clean_corpus <- function(corpus){
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, content_transformer(replace_symbol))
corpus <- tm_map(corpus, removeWords, c(stopwords("english"), "will", "can"))
# We could add more stop words as above
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, content_transformer(removeNumPunct))
return(corpus)
}
NYT_clean <- clean_corpus(NYT)
NYT_stem <- tm_map(NYT_clean, stemDocument)
meta(NYT_stem, type = "local", tag = "author") <- documents$author
NYT_dtm <- DocumentTermMatrix(NYT_stem)
NYT_tdm <- TermDocumentMatrix(NYT_stem)

NYT_tdm2 <- tidy(NYT_tdm)
NYT_dtm2 <- tidy(NYT_dtm)
NYT_tidy <- tidy(NYT_stem)
NYT_tdm2 <- merge(NYT_tdm2, NYT_tidy, by.x = "document", by.y = "id", all.x =
TRUE)
NYT_dtm2 <- merge(NYT_dtm2, NYT_tidy, by.x = "document", by.y = "id", all.x =
TRUE)

```


__Figure 12__

##### Wordcloud by Year

```{r, eval = TRUE, echo = FALSE, warning = FALSE, tidy = TRUE}
atxt <- documents$texts[documents$datetimestamp == "2011"]
btxt <- documents$texts[documents$datetimestamp == "2012"]
ctxt <- documents$texts[documents$datetimestamp == "2013"]
dtxt <- documents$texts[documents$datetimestamp == "2014"]
etxt <- documents$texts[documents$datetimestamp == "2015"]
ftxt <- documents$texts[documents$datetimestamp == "2016"]
gtxt <- documents$texts[documents$datetimestamp == "2017"]

clean.text <- function(x)
{
# tolower
x = tolower(x)
# remove rt
x = gsub("rt", "", x)
# remove at
x = gsub("@\\w+", "", x)
# remove punctuation
x = gsub("[[:punct:]]", "", x)
# remove numbers
x = gsub("[[:digit:]]", "", x)
# remove links http
x = gsub("http\\w+", "", x)
# remove tabs
x = gsub("[ |\t]{2,}", "", x)
# remove blank spaces at the beginning
x = gsub("^ ", "", x)
# remove blank spaces at the end
x = gsub(" $", "", x)
return(x)
}

aclean <- clean.text(atxt)
bclean <- clean.text(btxt)
cclean <- clean.text(ctxt)
dclean <- clean.text(dtxt)
eclean <- clean.text(etxt)
fclean <- clean.text(ftxt)
gclean <- clean.text(gtxt)

a <- paste(aclean, collapse=" ")
b <- paste(bclean, collapse=" ")
c <- paste(cclean, collapse=" ")
d <- paste(dclean, collapse=" ")
e <- paste(eclean, collapse=" ")
f <- paste(fclean, collapse=" ")
g <- paste(gclean, collapse=" ")

# put everything in a single vector
all <- c(a,b,c,d,e,f,g)
all <- removeWords(all, c("will", "can", "the", "that","are","mrs","not","said","cou"))
# create corpus
corpus2 <- Corpus(VectorSource(all))
# create term-document matrix
cloudtdm <- TermDocumentMatrix(corpus2)
# convert as matrix
cloudtdm <- as.matrix(cloudtdm)
# add column names
colnames(cloudtdm) <- c("2011","2012","2013","2014","2015","2016","2017")
comparison.cloud(cloudtdm, random.order=FALSE,colors = brewer.pal(8, "Dark2"),title.size=1.5, max.words = 200)
```
